Study of Higgs plus single top production using events with a same sign dimuon at the Large Hadron Collider

### 2019-11-14 presentacion_hiram/presentacion-tesis.tex
new

\documentclass[11pt]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{array}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{caption}
\usepackage{multicol}
\usepackage{listings}
\usepackage{alltt}
\usepackage{mathtools}
\usepackage[makeroom]{cancel}
\usetheme{Frankfurt}
\usepackage{tikz} 
\usetikzlibrary{patterns}
\usetikzlibrary{plotmarks}
\titlegraphic{\vspace{8cm}}% to push the other text to the top
\addtobeamertemplate{navigation symbols}{}{%
	\usebeamerfont{footline}%
	\usebeamercolor[fg]{footline}%
	\hspace{1em}%
	\insertframenumber/\inserttotalframenumber
	\setbeamercolor{footline}{fg=blue}
	\setbeamerfont{footline}{series=\bfseries}
}
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
		\newcommand\Fontvi{\fontsize{4}{7.2}\selectfont}
\begin{document}
	\author{Departamento de Investigacion en F\'isica \\
		Maestr\'ia en Ciencias (F\'isica)\\
		Hiram Ernesto Dami\'an\\
	Director:\\
Jos\'e Feliciano Ben\'itez Rubio}
	\title{Study of Higgs plus single top production using events with a same sign dimuon  at the Large Hadron Collider}
	%\subtitle{}
	\logo{\includegraphics[scale=0.1]{unison-logo.png}}
	\newcommand{\nologo}{\setbeamertemplate{logo}{}}
	\institute{Universidad de Sonora}
	%\date{}
	%\subject{}
	\newcommand{\subf}[2]{%
		{\small\begin{tabular}[t]{@{}c@{}}
				#1\\#2
		\end{tabular}}%
	}
	%\setbeamercovered{transparent}
	%\setbeamertemplate{navigation symbols}{}
\begin{frame}
\titlepage
\end{frame}

@@

\begin{frame}
\section{LHC and CMS}
\subsection{LHC}
\frametitle{LHC}
	\begin{figure}
		\centering
	\includegraphics[scale=0.2]{figures/CERN's-accelerator-complex2013.jpg}
	\end{figure}
\end{frame}

@@

\begin{frame}
\frametitle{LHC}
\begin{figure}[ht!]
	\centering
	\includegraphics[scale=0.6]{figures/csc.png}
	\caption*{Slide view of the CMS in terms of pseudorapidity $\eta=-\ln\left[\tan\left(\frac{\theta}{2}\right)\right]$}
\end{figure}
\end{frame}

@@

\begin{frame}
\subsection{BDT}
\frametitle{BDT}
\small{
	A decision tree takes a set of input features and splits input data recursively based on
	those features.
	Boosting is a method of combining many weak learners (trees) into a strong classifier. The features can be a mix of categorical and continuous data. \\

	BDT Variables}
\tiny{
	\begin{itemize}
		\item	Trailing lepton $p_{t}$
		\item 	Total charge of tight leptons
		\item 	min $\Delta$R (lepton pairs)
		\item 	$\Delta\phi$ between highest $p_t$ lepton pair
		\item 	Number of jets with |$\eta$| < 2.4
		\item	Number of non b-tagged jets with |$\eta$| > 1.0
		\item	Maximum |$\eta$| for jets
		\item	$\Delta\eta$ (most forward light jet, closest lepton)
		\item	$\Delta\eta$ (most forward light jet, hardest loosely b-tagged jet)
		\item	$\Delta\eta$ (most forward light jet, 2nd hardest loosely b-tagged jet)
\end{itemize} }
\end{frame}`

{\nologo
	\begin{frame}
	\Fontvi
	\begin{center}
		\begin{tabular}{cc}
			\includegraphics[width=5.5cm,height=3cm]{figures/distributions/compare_variables-abs-pj-eta} &
			\includegraphics[width=5.5cm,height=3cm]{figures/distributions/compare_variablescompare_variables-ncj24}\\ 
			{Distribution of selected events as a function of the jet with largest value of $|\eta|$} & {Distribution of selected events with higher pt with a |$\eta$| < 2.4 } \\
			\includegraphics[width=5.5cm,height=3cm]{figures/distributions/compare_variables-nfj}&
			\includegraphics[width=5.5cm,height=3cm]{figures/distributions/compare_variables-pt-sub}\\
			{Distribution of jet events for $|\eta|$>1 } & {Distribution of events of pt of the sub leading leptons} \\
		\end{tabular}
	\end{center}
\end{frame}
}

{\nologo
\begin{frame}
\Fontvi
\begin{center}
	\begin{tabular}{cc}
	\includegraphics[width=5.5cm,height=3cm]{figures/distributions/compare_variables-eta-fj-cl} &
\includegraphics[width=5.5cm,height=3cm]{figures/distributions/compare_variables-phi}\\ 
{Distribution of events of $|\Delta\eta|$ between a forward jet and the closest lepton} & {Distribution of events of $\Delta\phi$ between the higher pt pair of leptons } \\
		\includegraphics[width=5.5cm,height=3cm]{figures/distributions/compare_variables-dr}&
		\includegraphics[width=5.5cm,height=3cm]{figures/distributions/compare_variables-eta-fj-b1}\\
		{Distribution of events with a $\Delta R$ between the lepton pairs (highest and lowest pt)} & {Distribution of events as function of $|\Delta\eta|$ between a forward jet and a b-tagged jet } \\
	\end{tabular}
\end{center}
\end{frame}
}

@@

\begin{frame}
\frametitle{Results}
\framesubtitle{Likelihood scan}
{\small
	\begin{itemize}
\item Due to the large background, the signal strength for the Asimov data with 35.9 fb$^{-1}$ is consistent with zero.
Therefore, we estimate an upper limit on the signal strength at 95$\%$ confidence level.

\item We can define the likelihood ratio
\begin{align}
\lambda(\mu,\alpha)=\frac{L(\mu,\alpha)}{L(\hat{\mu},\hat{\alpha})}
\end{align}
Where $\hat{\alpha}$ and $\hat{\mu}$ are the parameters obtained in the previous section which correspond to the minimal of the NLL.

\item To determine an upper limit on the strength parameter $\mu$, we use the following statistical test
\begin{align}
q(\mu,\alpha)= -2\ln{\lambda} 
\end{align}
\item High values of $q$ represent greater incompatibility between the data and the fit model.
$q$ is a random variable with a $\chi^2$ distribution.
\end{itemize}
}
\end{frame}

@@

	\begin{thebibliography}{99}
		\bibliographystyle{unsrt}
	\begin{frame}
	\small
%	\section{References}
	\frametitle{References}
	\bibliographystyle{amsalpha}
	\bibitem{1}Gross F. \textit{Relativistic quantum mechanics and field theory}, 1994 ,WllEY-VCH Verlag GmbH \& Co. KGaA
	\bibitem{2} Griffiths, D. \textit{Introduction to Elementary Particles}, $2^\frac{\circ}{}$ edition, 2008,WllEY-VCH Verlag GmbH \& Co. KGaA
		\bibitem{2}The CMS collaboration, \textit{Search for production of a Higgs boson and a single top
	quark in multilepton final states in proton collisions at 13 TeV}, CMS-HIG-18-009, Phys. Rev. D 99, 092005
	\bibitem{3}Verkerke W \textit{Dealing with systematic uncertainties} 2014. From
\url{https://indico.cern.ch/event/287744/contributions/1641261/attachments/535763/738679/Verkerke_Statistics_3.pdf}
\end{frame}

\begin{frame}
\small
\bibitem{5} ATLAS and CMS
Collaborations, \textit{Combined measurement of the Higgs boson mass in
	pp collisions at $\sqrt{s}$= 7 and 8 TeV with the ATLAS and CMS experiments}, Phys. Rev. Lett.
114 (2015) doi:10.1103/PhysRevLett.114.191803, arXiv:1503.07589
\bibitem{4} Cowan G. , Cranmer K., Gross E. , Vitells O.\textit{ Asymptotic formulae for
	likelihood-based tests of new physics} 2013 , arXiv:1007.1727
\bibitem{7}CMS collaboration \textit{Search for tHq production in multilepton final states at 13 TeV} ,2017,CMS AN-16-378
\bibitem{6}CMS collaboration \textit{High-Luminosity Large Hadron Collider (HL-LHC)}, 2017,CERN-2017-007-M

\end{frame}
\end{thebibliography}

\begin{frame}
%\section{Supportive slides}
\huge{Back up}
\end{frame}

\begin{frame}
\frametitle{Sources of uncertainty on event yields}
\begin{itemize}
	\item Luminosity measurement: ~2.6$\%$.
	\item Data/MC scale factors for lepton selection (ID, iso) and trigger efficiencies ~5$\%$ per lepton.
	\item Choice of PDF set:
	\begin{itemize}
		\item 3.7$\%$ for tHq
		\item ~4$\%$ for tHW t$\bar{t}$W, t$\bar{t}$Z, t$\bar{t}$H 
		\item Scale uncertainties: 12$\%$, 4 for t$\bar{t}$W, 10$\%$ for t$\bar{t}$Z, +5.8/-9.2$\%$ for t$\bar{t}$H.
	\end{itemize}
	\item Background: WZ, ZZ sample modelling and statistics: ~50$\%$.
	\item Rare SM (tZ,tri-bosons, WWqq, tttt) : 50$\%$
	\item Fake rate estimation:
	The predicted event yield has a normalization uncertainty of ~30-50$\%$ \cite{7}
\end{itemize}

@@

\begin{frame}
\frametitle{Sources of systematic uncertainty}
Detector-simulation related uncertainty
\begin{itemize}
\item Calibrations (electron, jet energy scale)
\item Efficiencies (particle ID, reconstruction)
\item Resolutions (jet energy, muon momentum)
\end{itemize}
Theoretical uncertainties
\begin{itemize}
\item  Factorization/Normalization scale of MC generators
\item Choice of MC generator (ME and/or PS, e.g. Herwig vs Pythia)
\end{itemize}
Monte Carlo Statistical uncertainties
\begin{itemize}\item Statistical uncertainty of simulated samples\cite{2} \end{itemize}
\end{frame}

\begin{frame}
\begin{figure}
	\centering
	\includegraphics[scale=0.5]{figures/compare_variables}
	\caption*{Signal and background kinematic distributions}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Results}
\framesubtitle{Likelihood scan}
\begin{itemize}
	\item Likelihood function (often simply the likelihood) is a function of the
	parameters of a statistical model, given specific observed data.
	\item Likelihood functions play a key role in frequentist inference,
	especially methods of estimating a parameter from a set of
	statistics.
	\item In informal contexts, "likelihood" is often used as a synonym for
	probability.
\end{itemize}
\end{frame}

>>>>

\documentclass[11pt]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{array}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{caption}
\usepackage{multicol}
\usepackage{listings}
\usepackage{alltt}
\usepackage{mathtools}
\usepackage[makeroom]{cancel}
\usetheme{Frankfurt}
\usepackage{tikz} 
\usetikzlibrary{patterns}
\usetikzlibrary{plotmarks}
\titlegraphic{\vspace{8cm}}% to push the other text to the top
\addtobeamertemplate{navigation symbols}{}{%
	\usebeamerfont{footline}%
	\usebeamercolor[fg]{footline}%
	\hspace{1em}%
	\insertframenumber/\inserttotalframenumber
	\setbeamercolor{footline}{fg=blue}
	\setbeamerfont{footline}{series=\bfseries}
}
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
		\newcommand\Fontvi{\fontsize{4}{7.2}\selectfont}
\begin{document}
	\author{\tiny
			L.F. Hiram Ernesto Dami\'an\\
			\vspace{5px}
				Director:\\
					\vspace{5px}
			Jos\'e Feliciano Ben\'itez Rubio\\
				\vspace{5px}
				Maestr\'ia en Ciencias (F\'isica)\\
					\vspace{5px}
			 Departamento de Investigacion en F\'isica }
	\title{Study of Higgs plus single top production using events with a same sign dimuon  at the Large Hadron Collider}
	%\subtitle{}
	\logo{\includegraphics[scale=0.1]{unison-logo.png}}
	\newcommand{\nologo}{\setbeamertemplate{logo}{}}
	\institute{Universidad de Sonora}
	\date{\small November 15$^{th}$ 2019}
	%\subject{}
	\newcommand{\subf}[2]{%
		{\small\begin{tabular}[t]{@{}c@{}}
				#1\\#2
		\end{tabular}}%
	}
	%\setbeamercovered{transparent}
	%\setbeamertemplate{navigation symbols}{}
\begin{frame}
\titlepage
\end{frame}

@@

\begin{frame}
\section{LHC and CMS}
\subsection{LHC}
\frametitle{LHC}
	\begin{figure}
		\centering
	\includegraphics[scale=0.2]{figures/CERN's-accelerator-complex2013.jpg}
	\caption*{LHC complex}
	\end{figure}
\end{frame}

\begin{frame}
\frametitle{LHC}
\begin{figure}
	\centering
	\includegraphics[scale=0.6]{figures/lhc-u.png}
	\caption*{Underground view of LHC complex}
\end{figure}
\end{frame}

@@

\begin{frame}
\frametitle{CMS detector}
\begin{figure}[ht!]
	\centering
	\includegraphics[scale=0.6]{figures/csc.png}
	\caption*{Slide view of the CMS in terms of pseudorapidity $\eta=-\ln\left[\tan\left(\frac{\theta}{2}\right)\right]$}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{CMS detector}
\begin{figure}
	\centering
	\includegraphics[scale=0.4]{figures/pp.png}
	\caption*{\small A typical CMS event display of the Higgs boson decaying to four leptons with 2 muons (in red) and 2 electrons (in green) as final state signatures.}
\end{figure}
\end{frame}

@@

\begin{frame}
\subsection{BDT}
\frametitle{BDT}
\small{
	A decision tree takes a set of input features and splits input data recursively based on
	those features.
	Boosting is a method of combining many weak learners (trees) into a strong classifier. The features can be a mix of categorical and continuous data.\\
	\vspace{10px}
	BDT Variables}
\tiny{
	\begin{itemize}
		\item	Trailing lepton $p_{t}$
		\item 	Total charge of tight leptons
		\item 	min $\Delta$R (lepton pairs)
		\item 	$\Delta\phi$ between highest $p_t$ lepton pair
		\item 	Number of jets with |$\eta$| < 2.4
		\item	Number of non b-tagged jets with |$\eta$| > 1.0
		\item	Maximum |$\eta$| for jets
		\item	$\Delta\eta$ (most forward light jet, closest lepton)
		\item	$\Delta\eta$ (most forward light jet, hardest loosely b-tagged jet)
		\item	$\Delta\eta$ (most forward light jet, 2nd hardest loosely b-tagged jet)
\end{itemize} }
\end{frame}

{\nologo
	\begin{frame}
	\Fontvi
	\begin{center}
		\begin{tabular}{cc}
			\includegraphics[width=5.5cm,height=3cm]{figures/distributions/compare_variables-pt-sub} &
			\includegraphics[width=5.5cm,height=3cm]{figures/distributions/compare_variables-abs-pj-eta}\\ 
			  {Distribution of events of pt of the sub leading leptons} &{Distribution of selected events as a function of the jet with largest value of $|\eta|$}\\
			  	\includegraphics[width=5.5cm,height=3cm]{figures/distributions/compare_variablescompare_variables-ncj24}&
			\includegraphics[width=5.5cm,height=3cm]{figures/distributions/compare_variables-nfj}\\		
			 {Distribution of selected events with higher pt with a |$\eta$| < 2.4 } & {Distribution of jet events for $|\eta|$>1 } \\
		\end{tabular}
	\end{center}
\end{frame}
}

{\nologo
\begin{frame}
\Fontvi
\begin{center}
	\begin{tabular}{cc}
	\includegraphics[width=5.5cm,height=3cm]{figures/distributions/compare_variables-eta-fj-cl} &
\includegraphics[width=5.5cm,height=3cm]{figures/distributions/compare_variables-eta-fj-b1}\\ 
{Distribution of events of $|\Delta\eta|$ between a forward jet and the closest lepton} & {	Distribution of events as function of $|\Delta\eta|$ between a forward jet and a b-tagged jet } \\
		\includegraphics[width=5.5cm,height=3cm]{figures/distributions/compare_variables-dr}&
		\includegraphics[width=5.5cm,height=3cm]{figures/distributions/compare_variables-phi}\\
		{Distribution of events with a $\Delta R$ between the lepton pairs (highest and lowest pt)} & {Distribution of events of $\Delta\phi$ between the higher pt pair of leptons } \\
	\end{tabular}
\end{center}
\end{frame}
}

@@

\begin{frame}
\frametitle{Results}
\framesubtitle{Likelihood scan}
{\small
	\begin{itemize}
\item Due to the large background, the signal strength for the Asimov data with 35.9 fb$^{-1}$ is consistent with zero.
Therefore, we estimate an upper limit on the signal strength at 95$\%$ confidence level.
\item We can define the likelihood ratio
\begin{align}
\lambda(\mu,\alpha)=\frac{L(\mu,\alpha)}{L(\hat{\mu},\hat{\alpha})}
\end{align}
Where $\hat{\alpha}$ and $\hat{\mu}$ are the parameters obtained in the previous section which correspond to the minimal of the NLL.
\item To determine an upper limit on the strength parameter $\mu$, we use the following statistical test
\begin{align}
q(\mu,\alpha)= -2\ln{\lambda} 
\end{align}
\item High values of $q$ represent greater incompatibility between the data and the fit model.
$q$ is a random variable with a $\chi^2$ distribution.
\end{itemize}
}
\end{frame}

@@

	\begin{thebibliography}{99}
		\bibliographystyle{unsrt}
	\begin{frame}
	\tiny
%	\section{References}
	\frametitle{References}
	\bibliographystyle{amsalpha}
	\bibitem{1}Gross F. \textit{Relativistic quantum mechanics and field theory}, 1994 ,WllEY-VCH Verlag GmbH \& Co. KGaA
	\bibitem{2} Griffiths, D. \textit{Introduction to Elementary Particles}, $2^\frac{\circ}{}$ edition, 2008,WllEY-VCH Verlag GmbH \& Co. KGaA
		\bibitem{2}The CMS collaboration, \textit{Search for production of a Higgs boson and a single top
	quark in multilepton final states in proton collisions at 13 TeV}, CMS-HIG-18-009, Phys. Rev. D 99, 092005
	\bibitem{3}Verkerke W \textit{Dealing with systematic uncertainties} 2014. From
\url{https://indico.cern.ch/event/287744/contributions/1641261/attachments/535763/738679/Verkerke_Statistics_3.pdf}
\bibitem{5} ATLAS and CMS
Collaborations, \textit{Combined measurement of the Higgs boson mass in
	pp collisions at $\sqrt{s}$= 7 and 8 TeV with the ATLAS and CMS experiments}, Phys. Rev. Lett.
114 (2015) doi:10.1103/PhysRevLett.114.191803, arXiv:1503.07589
\bibitem{4} Cowan G. , Cranmer K., Gross E. , Vitells O.\textit{ Asymptotic formulae for
	likelihood-based tests of new physics} 2013 , arXiv:1007.1727
\bibitem{7}CMS collaboration \textit{Search for tHq production in multilepton final states at 13 TeV} ,2017,CMS AN-16-378
\bibitem{6}CMS collaboration \textit{High-Luminosity Large Hadron Collider (HL-LHC)}, 2017,CERN-2017-007-M
\end{frame}
\end{thebibliography}

%\begin{frame}
%\section{Supportive slides}
%\huge{Back up}
%\end{frame}

%\begin{frame}
%\frametitle{Sources of uncertainty on event yields}
%\begin{itemize}
%	\item Luminosity measurement: ~2.6$\%$.
%	\item Data/MC scale factors for lepton selection (ID, iso) and trigger efficiencies ~5$\%$ per lepton.
%	\item Choice of PDF set:
%	\begin{itemize}
%		\item 3.7$\%$ for tHq
%		\item ~4$\%$ for tHW t$\bar{t}$W, t$\bar{t}$Z, t$\bar{t}$H 
%		\item Scale uncertainties: 12$\%$, 4 for t$\bar{t}$W, 10$\%$ for t$\bar{t}$Z, +5.8/-9.2$\%$ for t$\bar{t}$H.
%	\end{itemize}
%	\item Background: WZ, ZZ sample modelling and statistics: ~50$\%$.
%	\item Rare SM (tZ,tri-bosons, WWqq, tttt) : 50$\%$
%	\item Fake rate estimation:
%	The predicted event yield has a normalization uncertainty of ~30-50$\%$ \cite{7}
%\end{itemize}
%\end{frame}

@@

%\begin{frame}
%\frametitle{Sources of systematic uncertainty}
%Detector-simulation related uncertainty
%\begin{itemize}
%\item Calibrations (electron, jet energy scale)
%%\item Efficiencies (particle ID, reconstruction)
%\item Resolutions (jet energy, muon momentum)
%\end{itemize}
%Theoretical uncertainties
%\begin{itemize}
%\item  Factorization/Normalization scale of MC generators
%\item Choice of MC generator (ME and/or PS, e.g. Herwig vs Pythia)
%\end{itemize}
%Monte Carlo Statistical uncertainties
%\begin{itemize}\item Statistical uncertainty of simulated samples\cite{2} \end{itemize}
%\end{frame}

%\begin{frame}
%\begin{figure}
%	\centering
%	\includegraphics[scale=0.5]{figures/compare_variables}
%	\caption*{Signal and background kinematic distributions}
%\end{figure}
%\end{frame}

%\begin{frame}
%\frametitle{Results}
%\framesubtitle{Likelihood scan}
%\begin{itemize}
%	\item Likelihood function (often simply the likelihood) is a function of the
%	parameters of a statistical model, given specific observed data.
%	\item Likelihood functions play a key role in frequentist inference,
%	especially methods of estimating a parameter from a set of
%	statistics.
%	\item In informal contexts, "likelihood" is often used as a synonym for
%	probability.
%\end{itemize}
%\end{frame}
